Stack Safety as a Security Property

### 2021-04-02 tex/main.tex
Finished up draft of examples section

@@

\begin{subfigure}{.25\textwidth}
\begin{verbatim}
int main() {
0[C]:  int t = f();
1[R]:  return;
}
\end{verbatim}
\end{subfigure}
\begin{subfigure}{.37\textwidth}
\begin{verbatim}
int f() {
10:     int x, y = 5, z = 6;
11:     print y+z;
12:     $SP-=2;
13[C]:  x = g();
14[R]:  return x;
}
\end{verbatim}
\end{subfigure}
\begin{subfigure}{.32\textwidth}
\begin{verbatim}
int g() {
20:     int z;
21[R]:  return (z+z);
}
\end{verbatim}
\end{subfigure}

\begin{subfigure}{.8\textwidth}  
\begin{center}
\begin{tabular}{l l l}
  {\tt 13} &
  \multicolumn{2}{c}{
    \memoryaddrs{16em}
    \memory{1}{\mainsealc}
    \memory{4}{\unsealc}
    ~$\cdots$
    \MemoryLabel{-23em}{0.75em}{a}
    \MemoryLabel{-19em}{0.75em}{b}
    \MemoryLabel{-15em}{0.75em}{1}
    \MemoryLabel{-11em}{0.75em}{5}
    \MemoryLabel{-8em}{0.75em}{6}
    \vspace{.5em}
  }
  &
\\
  {\tt 20} &
  \memoryaddrs{25em}
  \memory{1}{\mainsealc}
  \memory{2}{\fsealc}
  \memory{2}{\unsealc}
  ~$\cdots$
  \MemoryLabel{-23em}{0.75em}{a}
  \MemoryLabel{-19em}{0.75em}{b}
  \MemoryLabel{-15em}{0.75em}{1}
  \MemoryLabel{-11em}{0.75em}{14}
  \MemoryLabel{-8em}{0.75em}{\bf 6}
  \vspace{.5em}
  &
  \memoryaddrs{25em}
  \memory{1}{\mainsealc}
  \memory{2}{\fsealc}
  \memory{2}{\unsealc}
  ~$\cdots$
  \MemoryLabel{-23em}{0.75em}{a'}
  \MemoryLabel{-19em}{0.75em}{b'}
  \MemoryLabel{-15em}{0.75em}{c'}
  \MemoryLabel{-11em}{0.75em}{14}
  \MemoryLabel{-8em}{0.75em}{\bf d'}
  \vspace{.5em}
\\
\end{tabular}
\end{center}

@@

\begin{figure}
  \centering
  \begin{subfigure}[t]{.4\textwidth}
\begin{verbatim}
int stash = 0;

int main() {
0:     int x = 1;
1[C]:  f();
2:     x = -x;
3[C]:  f();
4[R]:  return x;
}

void f() {
10:     if(!stash) {
        // store return pointer
11:       stash = $SP-1;
12:     } else {
13:       $SP-1 = stash;
14:       stash = 0;
15[R]:    return;
        }
}
\end{verbatim}
  \end{subfigure}
\label{fig:WBCF1}
  \begin{subfigure}[t]{.4\textwidth}
\begin{verbatim}
int stash = 0;

int main() {
0:     int x = 1;
1[C]:  f();
2:     x = -x;
3[R]:  return x;
}

void f() {
10:     if(!stash) {
        // store return pointer
11:       stash = $SP-1;
12[C]     main();
12:     } else {
13:       $SP-1 = stash;
14:       stash = 0;
15[R]:    return;
        }
}
\end{verbatim}
\label{fig:WBCF2}
  \end{subfigure}
  \caption{Control Flow Attacks}
  \label{fig:WBCF}
\end{figure}

Example \ref{fig:WBCF1} shows how control flow attacks can compromise integrity in principle
even when the attacker obeys LSE. In this case, the attacker hides a return pointer in a
global variable and retrieves it the second time it is called, returning to the first call
point rather than the second. A return was performed, so {\tt x} is no longer sealed, but
line {\tt 2} is executed twice, so that {\tt main} returns 1 instead of -1.

@@

Non-skipping comes into play for attacks such as seen in \ref{WBCF2}, where the calls are
nested rather than sequential. \sna{The example isn't quite right; need to look at the
  Towards Complete... paper again.}

\sna{Put sharing example here, or after the formalism?}

>>>>

We alternatively could imagine a stepwise form of confidentiality in which we explicitly
track which addresses may be read (i.e., which have been initialized), and at each step
guarantee that the step alters state identically to a step from a variant state. Then we
wouldn't need to divide up the execution by calls. But this is further from the normal
framing of noninterference, requires explicitly tracking initializations, and would behave
no differently relative to our policies.

@@

\begin{subfigure}{.25\textwidth}
\begin{verbatim}
int main() {
0[C]:  int t = f();
1[R]:  return t;
}
\end{verbatim}
\end{subfigure}
\begin{subfigure}{.37\textwidth}
\begin{verbatim}
int f() {
10:     int x, y = 5, z = 6;
11:     print y+z;
12:     $SP-=2;
13[C]:  x = g();
14[R]:  return x;
}
\end{verbatim}
\end{subfigure}
\begin{subfigure}{.32\textwidth}
\begin{verbatim}
int g() {
20:     int z;
21[R]:  return (z+z);
}
\end{verbatim}
\end{subfigure}

\begin{subfigure}{.8\textwidth}  
\begin{center}
\begin{tabular}{l l l}
  {\tt 13} &
  \multicolumn{2}{c}{
    \memoryaddrs{16em}
    \memory{1}{\mainsealc}
    \memory{4}{\unsealc}
    ~$\cdots$
    \MemoryLabel{-23em}{0.75em}{a}
    \MemoryLabel{-19em}{0.75em}{b}
    \MemoryLabel{-15em}{0.75em}{1}
    \MemoryLabel{-11em}{0.75em}{5}
    \MemoryLabel{-8em}{0.75em}{6}
    \vspace{.5em}
  }
\\
  {\tt 20} &
  \memoryaddrs{25em}
  \memory{1}{\mainsealc}
  \memory{2}{\fsealc}
  \memory{2}{\unsealc}
  ~$\cdots$
  \MemoryLabel{-23em}{0.75em}{a}
  \MemoryLabel{-19em}{0.75em}{b}
  \MemoryLabel{-15em}{0.75em}{1}
  \MemoryLabel{-11em}{0.75em}{14}
  \MemoryLabel{-8em}{0.75em}{\bf 6}
  \vspace{.5em}
  &
  \memoryaddrs{25em}
  \memory{1}{\mainsealc}
  \memory{2}{\fsealc}
  \memory{2}{\unsealc}
  ~$\cdots$
  \MemoryLabel{-23em}{0.75em}{a'}
  \MemoryLabel{-19em}{0.75em}{b'}
  \MemoryLabel{-15em}{0.75em}{c'}
  \MemoryLabel{-11em}{0.75em}{14}
  \MemoryLabel{-8em}{0.75em}{\bf d'}
  \vspace{.5em}
\\
\end{tabular}
\end{center}

@@

\begin{figure}
  \centering
  \begin{minipage}{.4\textwidth}
\begin{verbatim}
int stash = 0;

int main() {
0:     int x = 1;
1[C]:  f();
2:     x = -x;
3[C]:  f();
4[R]:  return x;
}
\end{verbatim}
  \end{minipage}
  \begin{minipage}{.4\textwidth}
\begin{verbatim}
void f() {
10:     if(!stash) {
        // store return pointer
11:       stash = $SP-1;
12:     } else {
13:       $SP-1 = stash;
14:       stash = 0;
15[R]:    return;
        }
}
\end{verbatim}
  \end{minipage}
\label{fig:WBCF1}
%  \begin{subfigure}[t]{.4\textwidth}
%\begin{verbatim}
%int counter = 0;
%int stash = 0;
%
%int main() {
%0:     int x = 1;
%1[C]:  f();
%2:     x = -x;
%3[C]:  f();
%4[R]:  return x;
%}
%
%void f() {
%10:     switch(counter++) {
%11:       case 0:
%12[R]:      return;
%13:       case 1:
%14:         stash = $SP-1;
%15[C]:      main();
%16:         return;
%17:       default:
%18:         $SP-1 = stash;
%19[R]:      return;
%        }
%}
%\end{verbatim}
%\label{fig:WBCF2}
%  \end{subfigure}
  \caption{Control Flow Attacks}
  \label{fig:WBCF}
\end{figure}

Example \ref{fig:WBCF} shows how control flow attacks can compromise integrity in principle
even when the attacker obeys LSE. In this case, the attacker hides a return pointer in a
global variable and retrieves it the second time it is called, returning to the first call
point rather than the second. A return was performed, so {\tt x} is no longer sealed, but
line {\tt 2} is executed twice, so that {\tt main} returns 1 instead of -1.

@@

Non-skipping may seem strange -- surely the depth counter ought to decrease by one at a time
as part of the property definition, and the problem comes when the return is to a shallower
call's return pointer, violating return integrity? Yet some weaker enforcement mechanisms,
such as those in which returns are implemented as invocation of a Cheri object capability,
do allow skipping when adversarial code acts unreasonably by leaking its own return capability,
reasoning that this is acceptable as long as only the unreasonable function is skipped
(e.g. \citep{Skorstengaard+19b}). A variant of our model supports this weaker form of WBCF,
but with tag-based policies we can enforce the non-skipping version entirely.
\sna{Cite future work for details when the relevant section is finished.}

\paragraph*{Pass by Reference and Stack Arguments}

\newcommand{\mainpassc}{magenta}
\newcommand{\sharec}{lgray}

\begin{figure}[b]
  \centering
  \begin{subfigure}{.45\textwidth}
\begin{verbatim}
int stash;

int main() {
0:             int x = 42, y = 0;
1[P(x)],2[C]:  f(&x);
3:             x = 0;
4[P(y)],5[C]:  f(&y);
6[R]:          return 0;
}

void f(int* a) {
10:  if(!stash)
11:    stash = a;
12:  *stash = 5;
13:  return;
}
\end{verbatim}
\end{subfigure}
  \begin{subfigure}{.5\textwidth}
    \begin{tabular}{l l}
      {\tt 2} &
      \memoryaddrs{12em}
      \memory{1}{\mainpassc}[{\makebox[0pt]{Pass(0)}}]%
      \memory{2}{\unsealc}[{\makebox[0pt]{Unsealed}}]
      ~$\cdots$ \\
      {\tt 10} &
      \memoryaddrs{17em}
      \memory{1}{\mainpassc}[{\makebox[0pt]{Pass(0)}}]%
      \memory{1}{\mainsealc}[{\makebox[0pt]{Seal(0)}}]%
      \memory{2}{\unsealc}
      ~$\cdots$ \\
      {\tt 3} &
      \memoryaddrs{12em}
      \memory{3}{\unsealc}
      ~$\cdots$ \\
      {\tt 5} &
      \memoryaddrs{12em}
      \memory{1}{\unsealc}
      \memory{1}{\mainpassc}
      \memory{1}{\unsealc}
      ~$\cdots$ \\
      {\tt 10} &
      \memoryaddrs{17em}
      \memory{1}{\mainsealc}
      \memory{1}{\mainpassc}
      \memory{2}{\unsealc}
      ~$\cdots$ \\
\end{tabular}
\end{subfigure}

\label{fig:passing}
\end{figure}

Up til now we have assumed that function arguments and returns are passed solely in registers.
This is an oversimplication, which we will now correct. There are important stack safety concerns
when passing arguments on the stack, as seen in example \ref{fig:passing}. In this example,
{\tt main} passes a reference to {\tt x}, but then clears {\tt x} and instead passes one to
{\tt y}. Before each call, when the address of a variable is taken, we add a new annotation
\(P(x)\) or \(P(y)\) identifying the memory to be passed, and our stack model marks it in
magenta. Then at the call, the caller's frame that is not passed is sealed, as usual. Passed
memory also tracks which depth owns it, and it may only be accessed by the immediate callee
(at its depth plus one).

But because {\tt f} stashes the original pointer to {\tt x}, it will overwrite {\tt x} while
{\tt x} is sealed, an integrity violation. This fits with the intuition of passing by reference,
that the passed variable might be read or written during the call, not that it is permanently
open to be read or written by all subsequent calls. This also applies to passing large objects
by value in the stack: the caller allocates them, and the relevant instructions are annotated to
indicate passing.

We do not model cases where permanent sharing is desireable, because they are uninteresting
and hard to motivate: we would expect to treat such sharing as completely unprotected, and
perhaps attach a separate memory safety policy to add some protection to such variables.
